---
title: "Building an API Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Building an API Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(beekeeper)
```

The goal of beekeeper is to streamline and standardize the creation of high-quality API-wrapper packages.

The process requires an API that follows the [OpenAPI Specification](https://spec.openapis.org/oas/v3.0.0) (the "OAS").
Some APIs that follow the OAS can be found on [APIs.guru](https://apis.guru/).
We will use the [APIs.guru API](https://api.apis.guru/v2/openapi.yaml) (no authentication) and the [OpenFEC API](https://api.apis.guru/v2/specs/fec.gov/1.0/openapi.yaml) (authentication using an API key) as examples.

## Step 0: Create your package

Before you can use beekeeper, you need to create a basic package skeleton.
We recommend that you start with `usethis::create_package()` to streamline the basic package-creation steps.

You can then use beekeeper to get basic API calls working in that package through a two-step process.

## Step 1: Configure your package

If your target API follows the OAS, you can use your API's OpenAPI Document to configure your package.

```{r config-guru}
# Note: Running these commands will create or overwrite "_beekeeper.yml" in your
# working directory.
url("https://api.apis.guru/v2/openapi.yaml") |>
  use_beekeeper(
    api_abbr = "guru"
  )

# Or for the OpenFEC API
url("https://api.apis.guru/v2/specs/fec.gov/1.0/openapi.yaml") |>
  use_beekeeper(
    api_abbr = "fec"
  )
```

## Step 2: Generate your package skeleton

With a valid `_beekeeper.yml` file, you can generate most of the rest of the package.

```{r generate}
# Note: Running this command will create or overwrite files in your R and
# tests/testthat directories.
generate_pkg()
```

`generate_pkg()` creates files defining the package in the `R` directory, and tests in the `tests/testthat` directory.
If you do not already use `{testthat}` in your package, it will set up `{testthat}` for you.
It also adds any necessary dependencies to your `DESCRIPTION` file.
For example, beekeeper's sister package `{nectar}` handles calls to APIs and processing of responses.

## The generated package

Behind the scenes, `generate_pkg()` calls several helper functions to create the package files.

Note: The files shown here are formatted for readability. 
The actual files sometimes have different line wrapping and spacing.
We recommend that you use Posit's [Air](https://posit-dev.github.io/air/) to automatically style your code.
You will also want to review the generated code (especially documentation) to make sure it is accurate and matches your package's style.

### `generate_pkg_shared_params()`

To make parameter documentation consistent across functions, the generated package includes a file `R/000-shared.R` (generated with `generate_pkg_shared_params()`) that defines shared parameter documentation.

```{r fec_shared}
#' Parameters used in multiple functions
#'
#' Reused parameter definitions are gathered here for easier editing.
#'
#' @param max_reqs (length-1 `integer`) The maximum number of separate requests to
#'   perform. Passed on to [nectar::req_perform_opinionated()].
#' @param max_tries_per_req (length-1 `integer`) The maximum number of times to attempt
#'   each individual request. Passed on to [nectar::req_perform_opinionated()].
#' @param req (`httr2_request`) The request object to modify.
#' @param ... These dots are for future extensions and must be empty.
#' @param api_key An API key provided by the API provider. This key is not
#'   clearly documented in the API description. Check the API documentation for
#'   details.
#'
#' @name .shared-params
#' @keywords internal
NULL
```

You will likely want to update the documentation for authentication parameters (e.g. `api_key` in the example above) to provide more specific instructions for users of your package.
Updating the documentation once (in this file) updates the documentation for all functions that use these parameters.

The file generated for the APIs.guru API is largely the same, but it does not include the `api_key` parameter, since that API does not require authentication.

### `generate_pkg_req_prepare()`

A `R/010-prepare.R` file is generated with `generate_pkg_req_prepare()`.
This file defines a function that can be used to create a request object (`httr2::request()`) for the API.
We do not export this function by default.
Instead, it is called by each endpoint function (see below).

```{r guru_prepare}
#' Generate a request for the APIs.guru API
#'
#' Prepare a request for the APIs.guru API, using the opinionated framework
#' defined in [nectar::req_init()], [nectar::req_modify()],
#' [nectar::req_tidy_policy()], and [nectar::req_pagination_policy()].
#'
#' You may wish to export this function (if the API changes often or you do not
#' fully implement the API, for example).
#'
#' @inheritParams nectar::req_prepare
#' @inherit nectar::req_prepare return
#' @keywords internal
guru_req_prepare <- function(
  path,
  query = list(),
  body = NULL,
  header = list(),
  cookie = list(),
  method = NULL,
  tidy_policy = nectar::tidy_policy_unknown(),
  call = rlang::caller_env()
) {
  req <- nectar::req_prepare(
    "https://api.apis.guru/v2",
    path = path,
    query = query,
    body = body,
    header = header,
    cookie = cookie,
    method = method,
    tidy_policy = tidy_policy,
    call = call
  )

  return(req)
}
```

```{r fec_prepare}
#' Generate a request for the OpenFEC API
#'
#' Prepare a request for the OpenFEC API, using the opinionated framework
#' defined in [nectar::req_init()], [nectar::req_modify()],
#' [nectar::req_tidy_policy()], and [nectar::req_pagination_policy()].
#'
#' You may wish to export this function (if the API changes often or you do not
#' fully implement the API, for example).
#'
#' @inheritParams .shared-params
#' @inheritParams nectar::req_prepare
#' @inherit nectar::req_prepare return
#' @keywords internal
fec_req_prepare <- function(
  path,
  query = list(),
  body = NULL,
  header = list(),
  cookie = list(),
  method = NULL,
  tidy_policy = nectar::tidy_policy_unknown(),
  api_key = Sys.getenv("FEC_API_KEY"),
  call = rlang::caller_env()
) {
  req <- nectar::req_prepare(
    "https://api.open.fec.gov/v1",
    path = path,
    query = query,
    body = body,
    header = header,
    cookie = cookie,
    method = method,
    tidy_policy = tidy_policy,
    call = call
  )
  req <- .fec_req_auth(req, api_key = api_key)
  return(req)
}
```

Notice that the function includes the authentication parameters (e.g. `api_key` in the example above) and calls the authentication function defined in `R/020-auth.R` (see below).

`generate_pkg_req_prepare()` also generates a test file `tests/testthat/test-010-prepare.R` that tests the functionality of the prepare function (and sets up `{testthat}` if your package is not already using it).

```{r test-010-prepare}
test_that("Can prepare a request without errors", {
  test_result <- expect_no_error(guru_req_prepare("testing"))
  expect_s3_class(test_result, c("nectar_request", "httr2_request"))
  expect_named(
    test_result,
    c("url", "method", "headers", "body", "fields", "options", "policies")
  )
  expect_contains(
    names(test_result$policies),
    "resp_tidy"
  )
})
```

### `generate_pkg_auth()`

If the API definition includes a "security" component, it is used to define authentication functions for the API in defined in `R/020-auth.R` (with `generate_pkg_auth()`).
The initial version of this file works, but you may want to edit the automatic output.
In the case of the OpenFEC API, the description specifies three security schemes that overlap with one another: one that sets an `X-Api-Key` field in the header, and two that set an `api_key` in the query string.
The names of the generated functions are based on the names of the security schemes in the OpenAPI document.

```{r fec_auth}
# These functions were generated by the {beekeeper} package, based on
# components@security_schemes from the source API description. You may want to
# delete unused options. In addition, APIs often have additional security
# options that are not formally documented in the API description. For example,
# for any `location = query` `api_key` options, it might be possible to instead
# pass the same parameter in a header, possibly with a different name. Consult
# the text description of authentication in your API documentation.

.fec_req_auth <- function(req, api_key = NULL) {
  if (!is.null(api_key)) {
    req <- .fec_req_auth_api_key_header_auth(req, api_key)
  }
  if (!is.null(api_key)) {
    req <- .fec_req_auth_api_key_query_auth(req, api_key)
  }
  if (!is.null(api_key)) {
    req <- .fec_req_auth_api_key(req, api_key)
  }
  return(req)
}

# An API key provided by the API provider. This key is not clearly documented in the API description. Check the API documentation for details.
.fec_req_auth_api_key_header_auth <- function(req, api_key) {
  nectar::req_auth_api_key(
    req,
    location = "header",
    parameter_name = "X-Api-Key",
    api_key = api_key
  )
}

# An API key provided by the API provider. This key is not clearly documented in the API description. Check the API documentation for details.
.fec_req_auth_api_key_query_auth <- function(req, api_key) {
  nectar::req_auth_api_key(
    req,
    location = "query",
    parameter_name = "api_key",
    api_key = api_key
  )
}

# An API key provided by the API provider. This key is not clearly documented in the API description. Check the API documentation for details.
.fec_req_auth_api_key <- function(req, api_key) {
  nectar::req_auth_api_key(
    req,
    location = "query",
    parameter_name = "api_key",
    api_key = api_key
  )
}
```

For the real (in-progress) [`{fecapi}`](https://jonthegeek.github.io/fecapi/) package, I deleted the two `query` functions, since the `header` function is sufficient and slightly more secure.
I also renamed the `header` function from `.fec_req_auth_api_key_header_auth` to `.fec_req_auth_api_key_header` to remove the redundant `_auth`.

### `generate_pkg_paths()`

For each endpoint defined in the "paths" component of the API definition, `generate_pkg_paths()` creates a file in `R/` that defines a function to call that endpoint.
The functions are named based on the "tag" associated with the endpoint, and either the "operationId" (if available) or the HTTP method + path.

This is one such path file, generated as `R/paths-apis-get_api.R` for the APIs.guru API:

```{r guru-paths-apis-get_api}
# These functions were generated by the {beekeeper} package, based on the paths
# element from the source API description. You should carefully review these
# functions.

#' Retrieve one version of a particular API
#'
#' Returns the API entry for one specific version of an API where there is no
#' serviceName.
#'
#' @param provider (length-1 `character`)
#' @param api (length-1 `character`)
#' @inheritParams .shared-params
#'
#' @returns `guru_get_api()`: The API response.
#' @export
guru_get_api <- function(provider, api, max_reqs = Inf, max_tries_per_req = 3) {
  req <- req_guru_get_api(provider = provider, api = api)
  resps <- nectar::req_perform_opinionated(
    req,
    max_reqs = max_reqs,
    max_tries_per_req = max_tries_per_req
  )
  return(nectar::resp_parse(resps))
}

#' @rdname guru_get_api
#' @returns `req_guru_get_api()`: (`httr2_request`) A [httr2::request()] object.
req_guru_get_api <- function(provider, api) {
  provider <- stbl::to_chr_scalar(provider)
  api <- stbl::to_chr_scalar(api)
  guru_req_prepare(
    path = c("/specs/{provider}/{api}.json", provider = provider, api = api),
    method = "get"
  )
}
```

And this is `R/paths-audit-get_audit_case.R` for the OpenFEC API, the first of 91 generated path files:

```{r fec-paths-audit-get_audit_case}
# These functions were generated by the {beekeeper} package, based on the paths
# element from the source API description. You should carefully review these
# functions.

#' Get audit case
#'
#' This endpoint contains Final Audit Reports approved by the Commission since
#' inception. The search can be based on information about the audited committee
#' (Name, FEC ID Number, Type, Election Cycle) or the issues covered in the
#' report.
#'
#' @param audit_case_id (length-1 `list`, optional) Primary/foreign key for
#'   audit tables
#' @param cycle (length-1 `list`, optional) Filter records to only those that
#'   are applicable to a given two-year period. This cycle follows the
#'   traditional House election cycle and subdivides the presidential and Senate
#'   elections into comparable two-year blocks. The cycle begins with an odd
#'   year and is named for its ending, even year.
#' @param sub_category_id (length-1 `character`, optional) The finding id of an
#'   audit. Finding are a category of broader issues. Each category has an
#'   unique ID.
#' @param sort_nulls_last (length-1 `logical`, optional) Toggle that sorts null
#'   values last
#' @param sort_hide_null (length-1 `logical`, optional) Hide null values on
#'   sorted column(s).
#' @param min_election_cycle (length-1 `list`, optional) Filter records to only
#'   those that are applicable to a given two-year period. This cycle follows
#'   the traditional House election cycle and subdivides the presidential and
#'   Senate elections into comparable two-year blocks. The cycle begins with an
#'   odd year and is named for its ending, even year.
#' @param audit_id (length-1 `list`, optional) The audit issue. Each subcategory
#'   has an unique ID
#' @param q (length-1 `list`, optional) The name of the committee. If a
#'   committee changes its name, the most recent name will be shown. Committee
#'   names are not unique. Use committee_id for looking up records.
#' @param per_page (length-1 `list`, optional) The number of results returned
#'   per page. Defaults to 20.
#' @param max_election_cycle (length-1 `list`, optional) Filter records to only
#'   those that are applicable to a given two-year period. This cycle follows
#'   the traditional House election cycle and subdivides the presidential and
#'   Senate elections into comparable two-year blocks. The cycle begins with an
#'   odd year and is named for its ending, even year.
#' @param candidate_id (length-1 `list`, optional) A unique identifier assigned
#'   to each candidate registered with the FEC. If a person runs for several
#'   offices, that person will have separate candidate IDs for each office.
#' @param committee_type (length-1 `list`, optional) The one-letter type code of
#'   the organization: - C communication cost - D delegate - E electioneering
#'   communication - H House - I independent expenditure filer (not a committee)
#'   - N PAC - nonqualified - O independent expenditure-only (super PACs) - P
#'   presidential - Q PAC - qualified - S Senate - U single candidate
#'   independent expenditure - V PAC with non-contribution account, nonqualified
#'   - W PAC with non-contribution account, qualified - X party, nonqualified -
#'   Y party, qualified - Z national party non-federal account
#' @param qq (length-1 `list`, optional) Name of candidate running for office
#' @param page (length-1 `list`, optional) For paginating through results,
#'   starting at page 1
#' @param committee_id (length-1 `list`, optional) A unique identifier assigned
#'   to each committee or filer registered with the FEC. In general committee
#'   id's begin with the letter C which is followed by eight digits.
#' @param api_key (length-1 `character`) API key for https://api.data.gov. Get
#'   one at https://api.data.gov/signup.
#' @param committee_designation (length-1 `character`, optional) Type of
#'   committee: - H or S - Congressional - P - Presidential - X or Y or Z -
#'   Party - N or Q - PAC - I - Independent expenditure - O - Super PAC
#' @param primary_category_id (length-1 `character`, optional) Audit category ID
#'   (table PK)
#' @param sort_null_only (length-1 `logical`, optional) Toggle that filters out
#'   all rows having sort column that is non-null
#' @param sort (length-1 `list`, optional) Provide a field to sort by. Use `-`
#'   for descending order. ex: `-case_no`
#' @inheritParams .shared-params
#'
#' @returns `fec_get_audit_case()`: The API response.
#' @export
fec_get_audit_case <- function(
  audit_case_id,
  cycle,
  sub_category_id,
  sort_nulls_last,
  sort_hide_null,
  min_election_cycle,
  audit_id,
  q,
  per_page,
  max_election_cycle,
  candidate_id,
  committee_type,
  qq,
  page,
  committee_id,
  committee_designation,
  primary_category_id,
  sort_null_only,
  sort,
  api_key = Sys.getenv("FEC_API_KEY"),
  max_reqs = Inf,
  max_tries_per_req = 3
) {
  req <- req_fec_get_audit_case(
    audit_case_id = audit_case_id,
    cycle = cycle,
    sub_category_id = sub_category_id,
    sort_nulls_last = sort_nulls_last,
    sort_hide_null = sort_hide_null,
    min_election_cycle = min_election_cycle,
    audit_id = audit_id,
    q = q,
    per_page = per_page,
    max_election_cycle = max_election_cycle,
    candidate_id = candidate_id,
    committee_type = committee_type,
    qq = qq,
    page = page,
    committee_id = committee_id,
    committee_designation = committee_designation,
    primary_category_id = primary_category_id,
    sort_null_only = sort_null_only,
    sort = sort,
    api_key = api_key
  )
  resps <- nectar::req_perform_opinionated(
    req,
    max_reqs = max_reqs,
    max_tries_per_req = max_tries_per_req
  )
  return(nectar::resp_parse(resps))
}

#' @rdname fec_get_audit_case
#' @returns `req_fec_get_audit_case()`: (`httr2_request`) A [httr2::request()] object.
req_fec_get_audit_case <- function(
  audit_case_id,
  cycle,
  sub_category_id,
  sort_nulls_last,
  sort_hide_null,
  min_election_cycle,
  audit_id,
  q,
  per_page,
  max_election_cycle,
  candidate_id,
  committee_type,
  qq,
  page,
  committee_id,
  committee_designation,
  primary_category_id,
  sort_null_only,
  sort,
  api_key = Sys.getenv("FEC_API_KEY")
) {
  sub_category_id <- stbl::to_chr_scalar(sub_category_id)
  sort_nulls_last <- stbl::to_lgl_scalar(sort_nulls_last)
  sort_hide_null <- stbl::to_lgl_scalar(sort_hide_null)
  committee_designation <- stbl::to_chr_scalar(committee_designation)
  primary_category_id <- stbl::to_chr_scalar(primary_category_id)
  sort_null_only <- stbl::to_lgl_scalar(sort_null_only)
  fec_req_prepare(
    path = "/audit-case/",
    method = "get",
    api_key = api_key,
    query = list(
      audit_case_id = audit_case_id,
      cycle = cycle,
      sub_category_id = sub_category_id,
      sort_nulls_last = sort_nulls_last,
      sort_hide_null = sort_hide_null,
      min_election_cycle = min_election_cycle,
      audit_id = audit_id,
      q = q,
      per_page = per_page,
      max_election_cycle = max_election_cycle,
      candidate_id = candidate_id,
      committee_type = committee_type,
      qq = qq,
      page = page,
      committee_id = committee_id,
      committee_designation = committee_designation,
      primary_category_id = primary_category_id,
      sort_null_only = sort_null_only,
      sort = sort
    )
  )
}
```

You may want to edit the generated functions to provide more informative names and documentation.
For example, several parameters in the OpenFEC API definition have unclear types that result in `length-1 list` parameters in the generated functions.
You can edit the generated functions to provide more specific types and documentation for these parameters.

`generate_pkg_paths()` also generates a test file for each path in `tests/testthat/`.
Each file includes a basic snapshot test for one path, wrapped with `httptest2::with_mock_dir()` (so it only hits the API the first time you run it with a given set of arguments).

For example, this is the generated `tests/testthat/test-paths-audit-get_audit_case.R` for the OpenFEC API:

```{r fec-test-paths-audit}
# These tests were generated by the {beekeeper} package, based on the paths
# element from the source API description. You will likely need to supply
# arguments for the tests to succeed. We recommend expanding these tests to
# check for specific expectations, rather than simply using a snapshot.

httptest2::with_mock_dir("api/path/audit/fec_get_audit_case", {
  test_that("fec_get_audit_case() returns expected result", {
    expect_snapshot({
      test_result <- fec_get_audit_case(
        audit_case_id,
        cycle,
        sub_category_id,
        sort_nulls_last,
        sort_hide_null,
        min_election_cycle,
        audit_id,
        q,
        per_page,
        max_election_cycle,
        candidate_id,
        committee_type,
        qq,
        page,
        committee_id,
        committee_designation,
        primary_category_id,
        sort_null_only,
        sort
      )
      test_result
    })
  })
})
```

Currently, you will need to fill in the arguments for these tests to run successfully.

Finally, `generate_pkg_paths()` creates a file `tests/testthat/setup.R`, which sets up `{httptest2}` for testing.

```{r setup-httptest2}
library(httptest2)

# Don't repeat the base url in httptest2 recording paths. This helps avoid
# issues with long paths on Windows.
set_redactor(function(x) {
  gsub_response(
    x,
    nectar::url_path_append("https://api.open.fec.gov/v1", "/"),
    "",
    fixed = TRUE
  )
})
```

Specifically, `httptest2::set_redactor()` is configured with `httptest2::gsub_response()` to remove the base URL from the recording paths, to avoid issues with long paths on Windows.

## Next steps

The generated package is a starting point, but you will need to edit the generated files to finalize your package.
For example, you will need to update the tests to provide specific arguments.
You may also want to provide more specific documentation and types for parameters, or to add additional tests.

Since beekeeper is under active development, we encourage you to [report any issues you encounter with the generated code](https://github.com/api2r/beekeeper/issues), and to [contribute improvements to the package](https://beekeeper.api2r.org/CONTRIBUTING.html).
